A Comparison of Parallel Graph Processing Benchmarks

### 2017-02-17 papers/publication/paper-pollard.tex
Sam's update before merge. Tweaking some wording.

Moreover, the complexity of Graphalytics can obfuscate the true behavior of the program. By default, Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm in seconds. For example, one run of a single source shortest paths algorithm took 5.6 seconds. However, perusing the log files reveals a more complete picture: of these 5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures. Furthermore, other performance-relavant details, such as the number of iterations run for PageRank, are not easily available using Graphalytics.
%We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.
% The full logs:
%Finished setting ids, time: 0.157009
%Starting sort
%Finished sort, time: 1.39571
%Finished setting edge pointers, time: 0.000133
%Starting build_dcsc
%Finished build_dcsc, time: 0.235948
%Finished setting ids, time: 0.154693
%Starting sort
%Finished sort, time: 1.75676
%Finished setting edge pointers, time: 0.000142
%Starting build_dcsc
%Finished build_dcsc, time: 0.238386
%Completed reading A from memory in 4.089144 seconds.
%Completed reading A from file in 5.114708 seconds
%Completed 19 iterations
%Timing results:
%- load graph: 5.11472 sec
%- initialize engine: 2.88486e-05 sec
%- run algorithm: 0.212451 sec
%- print output: 0.0799532 sec
%- deinitialize engine: 2.88486e-05 sec
%17:29:05.575 [INFO ] Benchmarking algorithm "Single source shortest paths" on graph "dota-league" took 5955 ms.

@@

As with Graphalytics, the initial development effort is high but Granula paired with graphalytics allows automatic execution and compilation of performance results. Likewise, our approach provides automatic execution and performance analysis without requiring a performance and execution model for each system.

@@

Each experiment uses 32 roots per graph. As with the Graph500, each root is selected to have a degree greater than 1. In the case of PageRank, we simply run PageRank 32 times.

When possible, we measure the data set construction time as the time to translate from the unstructured file data in RAM to the graph representation on which the algorithm can be performed. This is not possible for PowerGraph and GraphBIG because they read in the input file and build a graph simultaneously. The Graph500 builds the graph once and performs the BFS on each root in turn, speeding up experimentation but only giving a single data point for the data set construction.

We plot many of the results as box plots with an implied 32 datapoints per box.

@@

We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented on all systems.

@@

Our work aims to add similar rigor to other graph algorithms by borrowing heavily from the Graph500 specification. The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the actual BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}. We run the BFS using 32 random roots with the exception of PowerGraph which doesn't provide an reference implementation of BFS in its toolkits.

@@

In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a single dataset. For an explanation of each algorithm used, see \cite{Iosup:2016:Graphalyticstech}. Table~\ref{tab:graphalytics} shows only a broad overview of the two systems: GraphBIG and PowerGraph. In general, we see GraphBIG is more performant. However, results such as these are preliminary and do not show the complete picture.

@@

Figures~\ref{fig:bfs-time} and \ref{fig:sssp-time} show more detail, with box plots for the runtime distributions. However, even these do not give a complete picture of performance. Both of these show GraphMat to not be highly performant. One explanation is GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs. There is less discrepancy between the runtimes of SSSP (between 0.1 and 1.7 seconds) compared to BFS (0.01 and 1.7 seconds) but GAP is the clear winner in both cases. The data structure construction times for GAP and GraphMat are consistent; in both cases the platforms create the same data structure for both algorithms.

These results are consistent with \cite{Sundaram:2015:GraphMat} which lists GraphMat as more performant than PowerGraph in SSSP.

@@

Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS. We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time and $n$ is the number of threads. A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}. These plots show in general poor scaling throughout, though GraphBIG has the best scaling. A potential explanation is GraphBIG is the slowest system so there is more room for improvement. Additionally, a scale of 20 generates a small graph by today's standards and thus the focus is on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
\begin{figure}
	\centering
	\begin{minipage}{0.9\linewidth}
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup.pdf}
	\end{minipage}
	\begin{minipage}{0.9\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss.pdf}
	\end{minipage}
	\caption{The top figure has both axes logarithmic. The exception is the extra space at 72 threads for readability. Linear speedup is. GraphMat dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
	\label{fig:bfs-scaling}
\end{figure}
\todo{Rerun GraphMat on scale 22 with 2 threads. Surprising that it slows down}

@@

\begin{table}
	\caption{The data are generated from a Kronecker graph with $2^{22}$ vertices. Sleeping Energy refers to the power (in Watts) consumed during the Unix \texttt{sleep} function, multiplied by time. Effectively, this measures the amount of energy that would have been consumed even if nothing was running. The increase over sleep is the ratio of the first and third columns. These are all averaged over the 32 roots.}
	\centering
	\begin{tabular}{l|r|r|r|r}
			&	GAP  &    Graph500 & GraphBIG & GraphMat \\ \hline
		Time (s) &  0.01636 & 0.01884 & 1.600 & 1.424 \\
		Average Power per Root (W) & 72.38 & 97.17 & 78.01 & 70.12 \\
		Energy per Root (J) &	1.184 & 1.830 & 112.213 & 111.104 \\
		Sleeping Energy (J) & 0.4046  & 0.4660 & 39.591 &  35.234 \\
		Increase over Sleep & 2.926 & 3.928 & 2.834 & 3.153
	\end{tabular}
	\label{tab:power}
\end{table}

@@

Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets and the most scalable. However, this is only for relatively small graphs; all the graphs used here had at most $2^{22}$ vertices. It should be noted that GAP is also the most recent of projects. Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite to as a good test of performance.

>>>>

Moreover, the complexity of Graphalytics can obfuscate the true behavior of the program. By default, Graphalytics generates an HTML report listing the runtimes for each dataset and each algorithm in seconds. For example, one run of a single source shortest paths algorithm took 5.96 seconds. However, perusing the log files reveals a more complete picture: of these 5.96 seconds, 0.21 was spent computing the shortest paths while the remaining time consisted of building the necessary data structures. Furthermore, other performance-relavant details, such as the number of iterations run for PageRank, are not easily available using Graphalytics.
%We use GraphMat, the single source shortest paths algorithm, the dota-league dataset from the Game Trace Archive\cite{Guo:2012:GTA}, and the default settings as an example here.
% The full logs:
%Finished setting ids, time: 0.157009
%Starting sort
%Finished sort, time: 1.39571
%Finished setting edge pointers, time: 0.000133
%Starting build_dcsc
%Finished build_dcsc, time: 0.235948
%Finished setting ids, time: 0.154693
%Starting sort
%Finished sort, time: 1.75676
%Finished setting edge pointers, time: 0.000142
%Starting build_dcsc
%Finished build_dcsc, time: 0.238386
%Completed reading A from memory in 4.089144 seconds.
%Completed reading A from file in 5.114708 seconds
%Completed 19 iterations
%Timing results:
%- load graph: 5.11472 sec
%- initialize engine: 2.88486e-05 sec
%- run algorithm: 0.212451 sec
%- print output: 0.0799532 sec
%- deinitialize engine: 2.88486e-05 sec
%17:29:05.575 [INFO ] Benchmarking algorithm "Single source shortest paths" on graph "dota-league" took 5955 ms.

@@

As with Graphalytics, the initial development effort is high but Granula paired with Graphalytics allows automatic execution and compilation of performance results. Likewise, our approach provides automatic execution and performance analysis but without requiring a performance and execution model for each system.

@@

Each experiment uses 32 roots per graph. As with the Graph500, each root is selected to have a degree greater than 1. For PageRank, we simply run the algorithm 32 times.

When possible, we measure the dataset construction time as the time to translate from the unstructured file data in RAM to the graph representation on which the algorithm can be performed. This is not possible for PowerGraph and GraphBIG because they read in the input file and build a graph simultaneously. The Graph500 builds the graph once and performs the BFS on each root in turn, speeding up experimentation but only giving a single data point for the dataset construction.

Any box plot with the exception of the Graph500 (instead labeled as solid lines rather than boxes) is a representation of 32 datapoints per box.

@@

We consider three algorithms: Breadth First Search (BFS), Single Source Shortest Paths (SSSP), and PageRank, though not all algorithms are implemented for all systems.

@@

Our work aims to add similar rigor to other graph algorithms by borrowing heavily from the Graph500 specification. The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an unsorted edge list stored in RAM and the actual BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}. We run the BFS using 32 random roots with the exception of PowerGraph which doesn't provide an reference implementation of BFS in its toolkits.

@@

In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on the Dota-League dataset. An explanation of each algorithm is given in \cite{Iosup:2016:Graphalyticstech}. Table~\ref{tab:graphalytics} shows only a broad overview of the two systems: GraphBIG and PowerGraph. In general, we see GraphBIG is more performant. However, results such as these are preliminary and do not show the complete picture.

@@

Figures~\ref{fig:bfs-time} and \ref{fig:sssp-time} show performance results for a Kronecker graph with scale $22$. The box plots give an idea of the runtime distributions. There is less discrepancy between the runtimes of SSSP (between 0.1 and 1.7 seconds) compared to BFS (0.01 and 1.7 seconds) but GAP is the clear winner in both cases. The data structure construction times for GAP and GraphMat are consistent; in both cases the platforms create the same data structure for both algorithms. These results are consistent with \cite{Sundaram:2015:GraphMat} which lists GraphMat as more performant than PowerGraph in SSSP.

@@

Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS. We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads. A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}. These plots show in general poor scaling throughout, though GraphBIG has the best scaling. A potential explanation is GraphBIG is the slowest system so there is more room for improvement. Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
\begin{figure}
	\centering
	\begin{minipage}{0.9\linewidth}
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup.pdf}
	\end{minipage}
	\begin{minipage}{0.9\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss.pdf}
	\end{minipage}
	\caption{The top figure has both axes logarithmic with an exception at 72 threads for readability. Linear speedup is. GraphMat dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
	\label{fig:bfs-scaling}
\end{figure}
\todo{Rerun GraphMat on scale 22 with 2 threads. Surprising that it slows down}

@@

\begin{table}
	\caption{The data are generated using a Kronecker graph with a scale of 22. Sleeping Energy refers to the power (in Watts) consumed during the \texttt{unistd} C \texttt{sleep} function, multiplied by the Time row. Essentially, this measures the amount of energy that would have been consumed even if nothing was running. The increase over sleep is the ratio of the first and third columns. These are all averaged over 32 roots.}
	\centering
	\begin{tabular}{l|r|r|r|r}
			&	GAP  &    Graph500 & GraphBIG & GraphMat \\ \hline
		Time (s) &  0.01636 & 0.01884 & 1.600 & 1.424 \\
		Average Power per Root (W) & 72.38 & 97.17 & 78.01 & 70.12 \\
		Energy per Root (J) &	1.184 & 1.830 & 112.213 & 111.104 \\
		Sleeping Energy (J) & 0.4046  & 0.4660 & 39.591 &  35.234 \\
		Increase over Sleep & 2.926 & 3.928 & 2.834 & 3.153
	\end{tabular}
	\label{tab:power}
\end{table}

@@

Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets but the least scalable. However, this is only for relatively small graphs; all the graphs used here had at most $2^{22}$ vertices. It should be noted that GAP is also the most recent of projects. Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite to as a good test of performance.

We found Graphmat to not be highly performant but relatively scalable. One explanation is GraphMat's underlying computation model (sparse matrix operations) paired the increased overhead of GraphMat's doubly-compressed sparse row (DSCR) graph representation is more conducive to larger-scale graphs.
% Furthermore, sparse matrix operations have been well studied so achieving good scalability is more realistic [cite]
